### 2021-01-07 paper/paper.tex
finish results, start rough draft

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Journal Article
% LaTeX Template
% Version 1.4 (15/5/16)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% Original author:
% Frits Wenneker (http://www.howtotex.com) with extensive modifications by
% Vel (vel@LaTeXTemplates.com)
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[twoside,twocolumn]{article}

\usepackage{blindtext} % Package to generate dummy text throughout this template

\usepackage[sc]{mathpazo} % Use the Palatino font
\usepackage[T1]{fontenc} % Use 8-bit encoding that has 256 glyphs
\linespread{1.05} % Line spacing - Palatino needs more space between lines
\usepackage{microtype} % Slightly tweak font spacing for aesthetics

\usepackage[english]{babel} % Language hyphenation and typographical rules

\usepackage[hmarginratio=1:1,top=32mm,columnsep=20pt]{geometry} % Document margins
\usepackage[hang, small,labelfont=bf,up,textfont=it,up]{caption} % Custom captions under/above floats in tables or figures
\usepackage{booktabs} % Horizontal rules in tables

\usepackage{lettrine} % The lettrine is the first enlarged letter at the beginning of the text

\usepackage{enumitem} % Customized lists
\setlist[itemize]{noitemsep} % Make itemize lists more compact

\usepackage{abstract} % Allows abstract customization
\renewcommand{\abstractnamefont}{\normalfont\bfseries} % Set the "Abstract" text to bold
\renewcommand{\abstracttextfont}{\normalfont\small} % Set the abstract itself to small italic text

\usepackage{titlesec} % Allows customization of titles
\renewcommand\thesection{\Roman{section}} % Roman numerals for the sections
\renewcommand\thesubsection{\Roman{subsection}} % roman numerals for subsections
\titleformat{\section}[block]{\large\scshape\centering}{\thesection.}{1em}{} % Change the look of the section titles
\titleformat{\subsection}[block]{\large}{\thesubsection.}{1em}{} % Change the look of the section titles

\usepackage{fancyhdr} % Headers and footers
\pagestyle{fancy} % All pages have headers and footers
\fancyhead{} % Blank out the default header
\fancyfoot{} % Blank out the default footer
\fancyhead[C]{Running title $\bullet$ May 2016 $\bullet$ Vol. XXI, No. 1} % Custom header text
\fancyfoot[RO,LE]{\thepage} % Custom footer text

\usepackage{titling} % Customizing the title section

\usepackage{hyperref} % For hyperlinks in the PDF

%----------------------------------------------------------------------------------------
%	TITLE SECTION
%----------------------------------------------------------------------------------------

\setlength{\droptitle}{-4\baselineskip} % Move the title up

\pretitle{\begin{center}\Huge\bfseries} % Article title formatting
\posttitle{\end{center}} % Article title closing formatting
\title{ISO-S: An ISODATA-Based Initial Centroid Algorithm} % Article title
\author{%
  \textsc{
    Noah Gardner$^1$\thanks{ngardn10@students.kennesaw.edu}, 
    Sahar Yarmohamadi$^2$\thanks{yarmohamadishr@gmail.com},
    Anthony Phan$^1$\thanks{aphan5@students.kennesaw.edu},
    Chih-Cheng Hung$^1$\thanks{chung1@students.kennesaw.edu}}\\[1ex]
  \normalsize $^1$Kennesaw State University \\
  \normalsize College of Computing and Software Engineering \\
  \normalsize Marietta, GA, USA
  \and
  \normalsize $^2$Institution 2 \\ %TODO 
}
\date{} % Leave empty to omit a date

\renewcommand{\maketitlehookd}{%
\begin{abstract}
\noindent 
Iterative Self-Organizing Data Analysis Technique (ISODATA) is a clustering
algorithm based on K-Means that intends to find the correct number of clusters
through merging and splitting. Compared to K-Means, the ISODATA algorithm should
be resilient to initial cluster centers. However, ISODATA is also sensitive to
initial cluster centers which impacts the performance of the algorithm. This
work proposes using the ISODATA approach of splitting clusters to select quality
initial cluster centers. The proposed approach is evaluated on several benchmark
UCI datasets as well as the Salinas-A hyperspectral image dataset. The results
are evaluated using standard clustering evaluation parameters. \\ \\
\textit{\textbf{Keywords:} Clustering, ISODATA, Initial Centroid, Remote Sensing} \\
\end{abstract}
}

%----------------------------------------------------------------------------------------

\begin{document}

% Print the title
\maketitle

%----------------------------------------------------------------------------------------
%	ARTICLE CONTENTS
%----------------------------------------------------------------------------------------

\section{Introduction}

\lettrine[nindent=0em,lines=3]{R} emote sensing technology is used when
the data collection takes place at a remote distance from the subject matter,
such as the study of daily weather and long-term climate change, land cover
monitoring, ecosystem modeling of vegetation, and so on. Multispectral imaging
technology is one of the instruments for remote sensing used since the late
1960s. However, multispectral imaging systems use a low number of spectral bands
(three to six spectral bands in different systems). Increasing spectral
resolution in the hyperspectral images (HSI) with several hundred spectral bands
of the observational scene in a single acquisition makes it possible to acquire
detailed examination of observational scenes. The ground-based version of
hyperspectral images has gained immense interest. It has been applied to
numerous electronic imaging applications, including food inspection, forensic
science, medical surgery and diagnosis, and military applications. Because of
the difficulty of obtaining prior knowledge in remote sensing studies,
unsupervised classification and cluster analysis have great importance compared
to supervised classification.

Unsupervised learning is a machine learning technique used when the learning
algorithm has no pre-existing labels and looks for patterns in the data set with
minimum human supervision and without external guidance. In contrast to
supervised learning, where a training set of inputs and observations are given
to the method, unsupervised learning uses only the inputs and should find an
interesting pattern in the data set. Two main approaches are used in
unsupervised learning, named principal component and cluster analysis. In the
clustering method, unlabeled and unsorted data points are grouped and segmented
according to similarities and differences to find structure in the inputs, group
it into meaningful clusters, and understand the data set better. There are
different algorithms for Clustering, such as K-Means, fuzzy C-Means, ISODATA,
and so on. In the K-Means algorithm, the goal is to minimize the sum of the
squared distances points have in the data set from its center. The problem with
the K-Means algorithm is that it requires a priori knowledge about the number of
classes in the data set, and choosing the cluster centers randomly may not
result in optimal partitioning of the data set.

One of the most popular clustering algorithms is called ISODATA clustering,
based on the K-means algorithm with some variations. The ISODATA method's
preference in comparison with the K-means algorithm is that the ISODATA
algorithm uses splitting and merging forms for clusters. The ISODATA method
allows starting from arbitrary cluster centers and obtains better partitioning
results than the K-Means algorithm, but it is not entirely resilient to initial
cluster centers. In this paper, a novel method using the ISODATA approach of
splitting clusters to select high-quality initial cluster centers is proposed.
This new method decreases the number of iterations and improves the overall
performance of the algorithm.

The paper is organized as follows: Section 2 introduces K-Means and ISODATA
clustering algorithms. The new proposed algorithm is explained in Section 3. We
present the experimental results in Section 4. Finally, Section 5 contains
concluding remarks.

%------------------------------------------------

\section{Methods}

\subsection{K-Means Algorithm}

K-Means algorithm is an unsupervised clustering algorithm and aims to partition
$N$ observations into $K$ clusters. Data points in the same cluster are very similar
to and with less variation. The user defines $K$, and these $K$ centers are placed
as far as possible for obtaining better clustering results. K-Means algorithm is
susceptible to initial starting cluster centers, and different initial values
result in a different classification. The next step is to assign each data point
to the nearest cluster center using Euclidean-based distance or other measures
based on the application-specific. After the assignment, the cluster centers are
updated by taking the average of all data points that belong to each cluster. We
continue this process until we meet the convergence criteria or the last
iteration. The algorithm is as follow:

\begin{itemize}
\item \textbf{Step 1:} Choose $K$ cluster centers from the data set.
\item \textbf{Step 2:} Assign each element of the data set to the cluster based
on the minimum distance from the cluster center.
\item \textbf{Step 3:} Re-calculate new cluster centers for each cluster after
the assignment in step 2, e.g., computing the mean of all patterns assigned to
each cluster.
\item \textbf{Step 4:} Continue steps 2 and 3 until Clustering convergence
occurs, or the number of iterations is exceeded.
\end{itemize}

For selecting the initial cluster centers in \textbf{Step 1}, there are two
standard methods:

\begin{itemize}
\item \textbf{Random Centers:} Explain here %TODO
\item \textbf{K-Means++:} Explain here %TODO
\end{itemize}

%------------------------------------------------

\subsection{ISODATA Algorithm}

ISODATA is another frequently used clustering algorithm and is similar to the
K-Means algorithm. However, in the ISODATA algorithm, the number of cluster
centers is determined dynamically, and it allows for a different number of
cluster centers. In contrast, the K-Means algorithm assumes that the number of
cluster centers is specified and fixated by the user. Some parameters such as
initial clusters' means, the minimum number of patterns in each cluster,
splitting and lumping parameters, and the number of iterations should be defined
and specified by the user. Like the K-Means algorithm, the first steps are
assignment all data points to the nearest cluster center and then recalculating
the cluster centers by replacing the average data point in each cluster. In the
next step, if a cluster does not reach the minimum number of clusters, the
user-specified, we eliminate it. After that, we do the splitting or lumping
process if their conditions are met. The cluster centers are updated based on
the newly grouped patterns. We continue this process until the convergence
occurs or the number of iterations is exceeded. The ISODATA algorithm is as
follow:

\begin{itemize}
\item \textbf{Step 1:} Choose $K$ cluster centers from the data set.
\item \textbf{Step 2:} Assign each element of the data set to the cluster based
on the minimum distance from the cluster center.
\item \textbf{Step 3:} Re-calculate new cluster centers for each cluster after
the assignment in step 2, e.g., computing the mean of all patterns assigned to
each cluster.
\item \textbf{Step 4:} Compare the number of patterns in each cluster against
the user's number ($\theta n$)and discard small sets. 
\item \textbf{Step 5:} Do splitting process based on splitting parameter
($\theta e$) if this iteration is an odd-numbered iteration of the number of
clusters is less than or equal to ½ the desired.
\item \textbf{Step 6:} Do merging process based on merging parameter ($\theta
c$)if this iteration is even-numbered, or if the number of clusters is greater
than or equal to twice the number desired.
\item \textbf{Step 7:} Start the process again at step 2 with new cluster
centers after splitting or lumping.
\item \textbf{Step 8:} If this is not the last iteration, start the process
again at step 2. If this is the last iteration, then end the process.
\end{itemize}

%------------------------------------------------

\subsection{Proposed Method: ISO-S}
The proposed clustering algorithm's objective is to improve the result we obtain
from the ISODATA algorithm by decreasing the number of iterations needed for
convergence. Inspired by the ISODATA algorithm, this capability is achieved
using the splitting technique to update each iteration cluster centers. The main
inputs for this algorithm are the arbitrary initial number of cluster centers
and the processing parameters. First, the algorithm starts with the one cluster
center and assign the patterns to this cluster center. Second, the new algorithm
updates the cluster center by computing the mean of data points. Finally, the
algorithm employs a splitting method to form new cluster centers. The cluster
centers with the smallest number of members are discarded at the end of the
maximum number of cluster centers we desire to obtain an odd number. We continue
this process until we meet the convergence criteria: the maximum number of
clusters desired or until we reach the maximum number of iterations. The
algorithm is outlined below:

%------------------------------------------------
\section{Results}

\blindtext % Dummy text

%------------------------------------------------

\section{Conclusion}

\subsection{Subsection One}

A statement requiring citation \cite{Figueredo:2009dg}.
\blindtext % Dummy text

\subsection{Subsection Two}

\blindtext % Dummy text

%----------------------------------------------------------------------------------------
%	REFERENCE LIST
%----------------------------------------------------------------------------------------

\begin{thebibliography}{99} % Bibliography - this is intentionally simple in this template

\bibitem[Figueredo and Wolf, 2009]{Figueredo:2009dg}
Figueredo, A.~J. and Wolf, P. S.~A. (2009).
\newblock Assortative pairing and life history strategy - a cross-cultural
  study.
\newblock {\em Human Nature}, 20:317--330.

\end{thebibliography}

%----------------------------------------------------------------------------------------

\end{document}

### 2021-01-07 ISODATA.notebook.jl

@@ -8,16 +8,20 @@ using InteractiveUtils
 begin
 	using Revise
 	
+	# dataset
+	using DataFrames
+	using CSV
+	
 	using BenchmarkTools
 	using Clustering
-	using CSV
-	using Distances
-	using DataFrames
 	using ISODATA
 	using MLDataUtils
 	using MLLabelUtils
 	using StatsBase
 end
+
+# ╔═╡ f277c7dc-4f64-11eb-147d-95a2b10bab9f
+
 
 # ╔═╡ c4a2ee6c-4f8d-11eb-24f5-e53ed7aa6789
 begin
@@ -33,6 +37,12 @@ begin
 	iris = cat(iX, transpose(iY), dims=1)
 end
 
+# ╔═╡ a1777880-5065-11eb-197d-0dba960318ed
+begin
+	si = summarize(iris)
+	si, iso_param(iris[1:end-1, :], si)
+end
+
 # ╔═╡ 7d36be2c-4f70-11eb-3b81-2b78885c3771
 begin
 	# load glass dataset
@@ -43,8 +53,18 @@ begin
 		MLLabelUtils.LabelEnc.Indices,
 		gY,
 		labelenc(gY))
+	for i = 5:7
+		inds = findall(x -> x == i, gY)
+		gY[inds] .= i - 1
+	end
 	gX = Array{Float64}(transpose(gX))
 	glass = cat(gX, transpose(gY), dims=1)
+end
+
+# ╔═╡ a3565518-5065-11eb-1b6f-27696c4519a6
+begin
+	sg = summarize(glass)
+	sg, iso_param(glass[1:end-1, :], sg)
 end
 
 # ╔═╡ b2ce85e2-4f8e-11eb-3c52-0fce3a112046
@@ -54,116 +74,6 @@ begin
 	wine_raw = convert(Matrix, wdf)
 	wX, wY = wine_raw[:, 2:end], wine_raw[:, 1]
 	wine = cat(transpose(wX), transpose(wY), dims=1)
-end
-
-# ╔═╡ affc89c4-4f64-11eb-29f8-15a7c437fe24
-struct Summary
-	label::Float64
-	means::Vector{Float64}
-	sd::Vector{Float64}
-	members::Int
-end
-
-# ╔═╡ f277c7dc-4f64-11eb-147d-95a2b10bab9f
-function summarize(data)
-	# assumes dimensions by samples format
-	d, n = size(data)
-	
[…]

### 2021-01-07 Salinas.notebook.jl

@@ -1,0 +1,332 @@
+### A Pluto.jl notebook ###
+# v0.12.18
+
+using Markdown
+using InteractiveUtils
+
+# ╔═╡ 0b8acfe4-5073-11eb-1630-33963891fa65
+begin
+	using Revise
+	
+	using MAT
+	
+	using Clustering
+	using ISODATA
+	using MLDataUtils
+	using MultivariateStats
+	using Plots
+	using StatsBase
+end
+
+# ╔═╡ 9d50e9dc-5112-11eb-261e-fb54e7238993
+gradient = cgrad([:black, :purple, :red, :yellow])
+
+# ╔═╡ 24a82cb0-5073-11eb-198e-a9de05e0c7a3
+begin
+	dfile = matopen("data/SalinasA.mat")
+	Xraw = read(dfile)["salinasA"]
+	close(dfile)
+	Xraw
+end
+
+# ╔═╡ 103c6df2-5077-11eb-1ee3-137d3269b6eb
+heatmap(Xraw[:, :, 4])
+
+# ╔═╡ a688fe58-5073-11eb-2968-bd2d796d1f2f
+begin
+	tfile = matopen("data/SalinasA_gt.mat")
+	Yraw = read(tfile)["salinasA_gt"]
+	close(tfile)
+	Yraw
+end
+
+# ╔═╡ b58883cc-5075-11eb-0c36-87c4d7f584f3
+heatmap(Yraw, c=gradient)
+
+# ╔═╡ 02200e30-5074-11eb-18c1-bbbd91537f47
+X = reshape(deepcopy(Xraw), 7138, 224)'
+
+# ╔═╡ 97b50c02-5074-11eb-1974-47e0028a67bb
+begin
+	Y = reshape(deepcopy(Yraw), 7138)
+	for i in 10:14
+		Y[findall(x -> x == Float64(i), Y)] .= i - 8.0
+	end
+	Y = round.(Int, Y)
+end
+
+# ╔═╡ efe37d6e-50f6-11eb-1e0d-c3b097dac6d6
+heatmap(reshape(Y, 83, 86), c=gradient)
+
+# ╔═╡ 59fa6a82-510d-11eb-06a9-c729dc0b2920
+struct ExperimentResult
+	iter::Float64
+	centers::Float64
+	adj_idx::Float64
+	rand_idx::Float64
+	mirkin_idx::Float64
+	hubert_idx::Float64
+end
+
[…]

### 2021-01-07 src/ISODATA.jl

@@ -9,10 +9,12 @@ include("centers.jl")
 include("clusters.jl")
 include("evaluation.jl")
 include("kmeans.jl")
+include("params.jl")
 
 export base, base_conv
 export randomcenters, kmppcenters, isodata_s
 export clusterfit
 export kmeans
+export summarize, iso_param
 
 end

### 2021-01-07 src/base.jl

@@ -138,7 +138,7 @@ function base_conv(
 
     while true
         iter += 1
-        if iter > 1000
+        if iter > 100
             return ClusteringResult(centers, assignments, costs, counts, sd, iter)
         end
 

### 2021-01-07 src/params.jl

@@ -1,0 +1,106 @@
+
+const _default_θn_bias = Float64(0.35)
+const _default_θe_bias = Float64(0.35)
+const _default_θc_bias = Float64(0.35)
+
+struct Summary
+    label::Float64
+    means::Vector{Float64}
+    sd::Vector{Float64}
+    members::Int
+end
+
+function summarize(data)
+    # assumes dimensions by samples format
+    d, n = size(data)
+
+    # assumes last row is class label
+    k = size(unique(data[end, :]), 1)
+
+    res = Vector{Summary}(undef, k)
+
+    for i in 1:k
+        inds = findall(x -> x == Float64(i), data[end, :])
+        temp = data[:, inds]
+
+        sd = Vector{Float64}(undef, d-1)
+        means = Vector{Float64}(undef, d-1)
+
+        for j in 1:d-1
+            means[j] = sum(temp[j, :])/size(temp, 2)
+            sd[j] = sqrt(sum((temp[j, :] .- means[j]).^2)/n)
+        end
+
+        res[i] = Summary(Float64(i), means, sd, size(temp, 2))
+    end
+
+    return res
+end
+
+"""
+    iso_param(data, summaries) -> Float64, Float64, Float64
+
+    Given data and the summarized data (based on class label),
+    attempt to generate ISODATA parameters based on a bias 
+    percentage.
+"""
+function iso_param(
+	data,
+	summaries::Array{Summary};
+	_θn_bias::Float64=_default_θn_bias,
+	_θe_bias::Float64=_default_θe_bias,
+	_θc_bias::Float64=_default_θc_bias
+)
+	d, n = size(data)
+	m = size(summaries, 1)
+	
+	θn, θe, θc = 0.0, 0.0, 0.0
+	
+	# find θn, minimum number of cluster members
+	min_members = 0 
+	total_members = 0
+	for i = 1:m
+		if min_members < summaries[i].members
+			min_members = summaries[i].members
+		end
+		total_members += summaries[i].members
+	end
+	
+	# θn is _θn_bias% lower than min_members/total_members
+	θn = (min_members/total_members)*(1-_θn_bias)
+	
+	# find θe, maximum standard deviation allowed for a cluster
+	max_sd = 0.0
+	for i = 1:m
+		for j = 1:d
+			if summaries[i].sd[j] > max_sd
+				max_sd = summaries[i].sd[j]
[…]

### 2021-01-08 paper/paper.tex
adjust algorithm details, finish proposed method section, start results

@@
 \posttitle{\end{center}} % Article title closing formatting
-\title{ISO-S: An ISODATA-Based Initial Centroid Algorithm} % Article title
+\title{ISODS: An ISODATA-Based Initial Centroid Algorithm} % Article title
 \author{%
@@
     Noah Gardner$^1$\thanks{ngardn10@students.kennesaw.edu}, 
-    Sahar Yarmohamadi$^2$\thanks{yarmohamadishr@gmail.com},
+    Sahar Yarmohammadtoosky$^2$\thanks{yarmohamadishr@gmail.com},
+    }\\[1ex]
+  \textsc{
     Anthony Phan$^1$\thanks{aphan5@students.kennesaw.edu},
-    Chih-Cheng Hung$^1$\thanks{chung1@students.kennesaw.edu}}\\[1ex]
+    Chih-Cheng Hung$^1$\thanks{chung1@students.kennesaw.edu}} \\[1ex]
   \normalsize $^1$Kennesaw State University \\
   \normalsize College of Computing and Software Engineering \\
-  \normalsize Marietta, GA, USA
-  \and
-  \normalsize $^2$Institution 2 \\ %TODO 
+  \normalsize Marietta, GA, USA \\[1ex]
+  \normalsize $^2$Iran University of Science and Technology \\
+  \normalsize Tehran, Iran
 }
@@
 
-\section{Methods}
-
 \subsection{K-Means Algorithm}
 
-K-Means algorithm is an unsupervised clustering algorithm and aims to partition
-$N$ observations into $K$ clusters. Data points in the same cluster are very similar
-to and with less variation. The user defines $K$, and these $K$ centers are placed
-as far as possible for obtaining better clustering results. K-Means algorithm is
-susceptible to initial starting cluster centers, and different initial values
-result in a different classification. The next step is to assign each data point
-to the nearest cluster center using Euclidean-based distance or other measures
-based on the application-specific. After the assignment, the cluster centers are
-updated by taking the average of all data points that belong to each cluster. We
-continue this process until we meet the convergence criteria or the last
-iteration. The algorithm is as follow:
+The K-Means algorithm is an unsupervised clustering algorithm and aims to
+partition $N$ observations into $K$ clusters [CITE]. Data points in the same
+cluster are very similar to each other and have less variation compared to data
+points assigned to other clusters. The user defines $K$, the number of expected
+clusters. Following is a description of the K-Means algorithm:
 
@@
 \item \textbf{Step 3:} Re-calculate new cluster centers for each cluster after
-the assignment in step 2, e.g., computing the mean of all patterns assigned to
-each cluster.
-\item \textbf{Step 4:} Continue steps 2 and 3 until Clustering convergence
-occurs, or the number of iterations is exceeded.
+the assignment in \textbf{Step 2} by computing the mean of all patterns
+assigned to each cluster.
+\item \textbf{Step 4:} Repeat \textbf{Step 2 and Step 3} until clustering
+convergence occurs, or the set number of iterations is exceeded.
 \end{itemize}
@@
 For selecting the initial cluster centers in \textbf{Step 1}, there are two
-standard methods:
-
-\begin{itemize}
-\item \textbf{Random Centers:} Explain here %TODO
-\item \textbf{K-Means++:} Explain here %TODO
-\end{itemize}
+standard methods: Random Centers, and K-Means++ [CITE]. In Random Centers, $K$
+samples are randomly chosen from the dataset to serve as the initial cluster
+centers. In K-Means++, $1$ sample is randomly chosen from the dataset to serve
+as the initial cluster center. Then, $K-1$ samples are iteratively selected,
+where the probability of a sample being chosen as a cluster center is the
+distance of the sample from previously chosen centers [CITE]. The K-Means
+algorithm is susceptible to initial starting cluster centers, where different
+initial values can result in a different classification.
 
@@
 
-ISODATA is another frequently used clustering algorithm and is similar to the
-K-Means algorithm. However, in the ISODATA algorithm, the number of cluster
-centers is determined dynamically, and it allows for a different number of
-cluster centers. In contrast, the K-Means algorithm assumes that the number of
-cluster centers is specified and fixated by the user. Some parameters such as
-initial clusters' means, the minimum number of patterns in each cluster,
-splitting and lumping parameters, and the number of iterations should be defined
-and specified by the user. Like the K-Means algorithm, the first steps are
-assignment all data points to the nearest cluster center and then recalculating
-the cluster centers by replacing the average data point in each cluster. In the
-next step, if a cluster does not reach the minimum number of clusters, the
-user-specified, we eliminate it. After that, we do the splitting or lumping
-process if their conditions are met. The cluster centers are updated based on
-the newly grouped patterns. We continue this process until the convergence
-occurs or the number of iterations is exceeded. The ISODATA algorithm is as
-follow:
+ISODATA is another frequently used clustering algorithm and is based on the
+K-Means algorithm. Like K-Means, cluster centers are iteratively updated and
+reassigned to the samples in the dataset. However, ISODATA determines the number
+of clusters dynamically and allows for a different number of cluster centers. In
+contrast, the K-Means algorithm requires a fixed user-set parameter of the
+number of expected clusters. The advantage found within ISODATA allows for
+discovery of a different number of clusters, which is achieved through
+discarding clusters with a small number of member samples, and merging and
+splitting clusters that meet certain parameters. There are three main ISODATA
+parameters:
+
+\begin{itemize}
+\item $\theta n$ - minimum number of cluster members, which determines whether
+the cluster is discarded,
+\item $\theta e$ - maximum standard deviation allowed for a cluster, which
+determines whether the cluster is split,
+\item $\theta c$ - minimum distance between clusters, which determines whether
+two clusters are merged.
+\end{itemize}
+
+Additionally, the user must set $K$, the maximum number of clusters to generate.
+The ISODATA algorithm is briefly explained below:
 
@@
 \item \textbf{Step 1:} Choose $K$ cluster centers from the data set.
-\item \textbf{Step 2:} Assign each element of the data set to the cluster based
+\item \textbf{Step 2:} Assign each element of the dataset to the cluster based
 on the minimum distance from the cluster center.
-\item \textbf{Step 3:} Re-calculate new cluster centers for each cluster after
-the assignment in step 2, e.g., computing the mean of all patterns assigned to
-each cluster.
-\item \textbf{Step 4:} Compare the number of patterns in each cluster against
-the user's number ($\theta n$)and discard small sets. 
-\item \textbf{Step 5:} Do splitting process based on splitting parameter
-($\theta e$) if this iteration is an odd-numbered iteration of the number of
-clusters is less than or equal to ½ the desired.
-\item \textbf{Step 6:} Do merging process based on merging parameter ($\theta
-c$)if this iteration is even-numbered, or if the number of clusters is greater
-than or equal to twice the number desired.
-\item \textbf{Step 7:} Start the process again at step 2 with new cluster
-centers after splitting or lumping.
-\item \textbf{Step 8:} If this is not the last iteration, start the process
-again at step 2. If this is the last iteration, then end the process.
+\item \textbf{Step 3:} Re-calculate new cluster centers for each cluster by
+computing the mean of all patterns assigned to each cluster.
+\item \textbf{Step 4:} Discard clusters that have fewer than ($\theta n$)
+members. If any clusters were discarded, move to \textbf{Step 7}.
+\item \textbf{Step 5:} If this iteration is an odd numbered iteration or the
+number of clusters is less than or equal to  $1/2 * K$, then split clusters
+which have a standard deviation in a dimension greater than ($\theta e$). Move
+to \textbf{Step 7}.
+\item \textbf{Step 6:} If this iteration is an even numbered iteration or the
+number of clusters is greater than or equal to $2 * K$, then merge clusters which
+have an intercluster distance less than ($\theta c$).
+\item \textbf{Step 7:} If this is not the last iteration, start the process
+again at step 2. If this is the last iteration or clustering convergence occurs,
+then end the process.
 \end{itemize}
+
+There are some disadvantages to the ISODATA algorithm. First, ISODATA is
+sensitive to the parameters $\theta n$, $\theta e$, and $\theta c$. These
+parameters should be fine-tuned for optimal performance based on the attributes
+of the dataset being tested. In addition, similar to K-Means, it is sensitive to
+initial cluster centers, although less so than K-Means due to the extra
+processes of discarding, splitting, and merging. The initial centroids can
+affect both the quality of clusters produced and the number of iterations
+required to converge [CITE].
 
@@
 
-\subsection{Proposed Method: ISO-S}
+\section{Proposed Method}
+
 The proposed clustering algorithm's objective is to improve the result we obtain
-from the ISODATA algorithm by decreasing the number of iterations needed for
-convergence. Inspired by the ISODATA algorithm, this capability is achieved
-using the splitting technique to update each iteration cluster centers. The main
-inputs for this algorithm are the arbitrary initial number of cluster centers
-and the processing parameters. First, the algorithm starts with the one cluster
-center and assign the patterns to this cluster center. Second, the new algorithm
-updates the cluster center by computing the mean of data points. Finally, the
-algorithm employs a splitting method to form new cluster centers. The cluster
-centers with the smallest number of members are discarded at the end of the
-maximum number of cluster centers we desire to obtain an odd number. We continue
-this process until we meet the convergence criteria: the maximum number of
-clusters desired or until we reach the maximum number of iterations. The
-algorithm is outlined below:
+from the ISODATA algorithm by targeting the initial cluster centers. Inspired by
+the ISODATA algorithm, Iterative Self-Organizing Data Split (ISODS) uses the
+splitting process from ISODATA to create the required initial cluster centers to
+use for the ISODATA algorithm. The inputs for the algorithm are the number of
+cluster centers $K$ as an integer, and the data to generate centers from. The
+algorithm starts by selecting $1$ sample as an initial cluster center. In the
+next step, each data is assigned to the cluster and the mean of the cluster is
+updated. Then, by the ISODATA splitting process, each cluster is split along the
+dimension with the highest standard deviation. The assignments for the new
+clusters are calculated based on distance and the centers are updated. This
+process is continued until either the number of clusters is greater than or
+equal to $K$. If the number of clusters is greater than $K$, the clusters with
+the fewest members are discarded one at a time, and the assignments and centers
+are updated after each rejection. The algorithm is outlined below:
+
+\begin{itemize}
+\item \textbf{Step 1:} Choose $1$ observation as a cluster center from the dataset.
+\item \textbf{Step 2:} Assign each element of the dataset to the cluster based
+on the minimum distance from the cluster center.
+\item \textbf{Step 3:} Re-calculate new cluster centers for each cluster by
+computing the mean of all patterns assigned to each cluster.
+\item \textbf{Step 4:} If the number of clusters is equal to $K$, then the
+process is terminated. If the number of clusters is greater than $K$, then the
+clusters with the fewest members is discarded, and the algorithm moves to
+\textbf{Step 2}.
+\item \textbf{Step 5:} Calculate the standard deviation of each cluster in each
+dimension. Then, split each cluster along the dimension with the greatest
+standard deviation.
+\end{itemize}
+
+This method began as a modified ISODATA process. However, the modification was
+unique in that there is a clear distinction from the proposed change and the
+beginning of the original process. Although more complex than other methods of
+generating initial cluster centers, this approach guarantees variance between
+clusters. As a result, ISODS can be used to generate clusters for other
+clustering algorithms, such as K-Means. In the next section, ISODS is applied to
+both ISODATA and K-Means on some benchmark UCI datasets [CITE] and compared with
+Random Centers and K-Means++ Centers.
 
@@
 
-\blindtext % Dummy text
+Due to the response of the ISODATA algorithm to the discarding, merging, and
+splitting parameters, these parameters should be carefully selected. Since the
+goal of this research is to explore the impact of initial cluster centers on
+ISODATA, we use a relatively simple approach for generating ISODATA parameters.
+Based on the available training data with included labels, each class is
+summarized with the mean and standard deviation of that class, as well as the
+number of members in each class. To determine $\theta n$, the minimum number of
+cluster members used for discarding, we find the class with the fewest members
+and use the number of members in that class as $\theta n$. For $\theta e$, the
+maximum standard deviation allowed for a cluster, we follow a similar approach
+and find the maximum standard deviation within a class. Finally, for $\theta c$,
+we calculate the distance between the means of each class, and use the minimum
+distance as $\theta c$.
 

### 2021-01-08 paper/paper.tex
add references

@@
 
-\lettrine[nindent=0em,lines=3]{R} emote sensing technology is used when
-the data collection takes place at a remote distance from the subject matter,
-such as the study of daily weather and long-term climate change, land cover
-monitoring, ecosystem modeling of vegetation, and so on. Multispectral imaging
-technology is one of the instruments for remote sensing used since the late
-1960s. However, multispectral imaging systems use a low number of spectral bands
-(three to six spectral bands in different systems). Increasing spectral
-resolution in the hyperspectral images (HSI) with several hundred spectral bands
-of the observational scene in a single acquisition makes it possible to acquire
-detailed examination of observational scenes. The ground-based version of
-hyperspectral images has gained immense interest. It has been applied to
-numerous electronic imaging applications, including food inspection, forensic
-science, medical surgery and diagnosis, and military applications. Because of
-the difficulty of obtaining prior knowledge in remote sensing studies,
-unsupervised classification and cluster analysis have great importance compared
-to supervised classification.
+\lettrine[nindent=0em,lines=3]{R} emote sensing technology is used when the data
+collection takes place at a remote distance from the subject matter, such as the
+study of daily weather and long-term climate change, land cover monitoring,
+ecosystem modeling of vegetation, and so on \cite{Jensen1996,Mallapragada2018}.
+Multispectral imaging technology is one of the instruments for remote sensing
+used since the late 1960s \cite{Khan2018,Landgrebe1999}. However, multispectral
+imaging systems use a low number of spectral bands (three to six spectral bands
+in different systems). Increasing spectral resolution in the hyperspectral
+images (HSI) with several hundred spectral bands of the observational scene in a
+single acquisition makes it possible to acquire detailed examination of
+observational scenes \cite{Khan2018}. The ground-based version of hyperspectral
+images has gained immense interest. It has been applied to numerous electronic
+imaging applications, including food inspection, forensic science, medical
+surgery and diagnosis, and military applications
+\cite{Govender2009,Adam2009,Fischer2006,Carrasco2003}. Because of the difficulty
+of obtaining prior knowledge in remote sensing studies, unsupervised
+classification and cluster analysis have great importance compared to supervised
+classification \cite{Abbas2016}.
 
@@
 interesting pattern in the data set. Two main approaches are used in
-unsupervised learning, named principal component and cluster analysis. In the
-clustering method, unlabeled and unsorted data points are grouped and segmented
-according to similarities and differences to find structure in the inputs, group
-it into meaningful clusters, and understand the data set better. There are
-different algorithms for Clustering, such as K-Means, fuzzy C-Means, ISODATA,
-and so on. In the K-Means algorithm, the goal is to minimize the sum of the
-squared distances points have in the data set from its center. The problem with
-the K-Means algorithm is that it requires a priori knowledge about the number of
-classes in the data set, and choosing the cluster centers randomly may not
-result in optimal partitioning of the data set.
+unsupervised learning, named principal component \cite{Mallapragada2018} and
+cluster analysis \cite{Xu2015}. In the clustering method, unlabeled and unsorted
+data points are grouped and segmented according to similarities and differences
+to find structure in the inputs, group it into meaningful clusters, and
+understand the data set better. There are different algorithms for Clustering,
+such as K-Means, fuzzy C-Means, ISODATA \cite{Ball1965}, and so on. In the
+K-Means algorithm, the goal is to minimize the sum of the squared distances
+points have in the data set from its center. The problem with the K-Means
+algorithm is that it requires a priori knowledge about the number of classes in
+the data set, and choosing the cluster centers randomly may not result in
+optimal partitioning of the data set.
 
@@
 cluster centers. In this paper, a novel method using the ISODATA approach of
-splitting clusters to select high-quality initial cluster centers is proposed.
-This new method decreases the number of iterations and improves the overall
-performance of the algorithm.
+splitting clusters to select quality initial cluster centers is proposed.
 
-The paper is organized as follows: Section 2 introduces K-Means and ISODATA
-clustering algorithms. The new proposed algorithm is explained in Section 3. We
-present the experimental results in Section 4. Finally, Section 5 contains
-concluding remarks.
+% The paper is organized as follows: Section 2 introduces K-Means and ISODATA
+% clustering algorithms. The new proposed algorithm is explained in Section 3. We
+% present the experimental results in Section 4. Finally, Section 5 contains
+% concluding remarks.
 
@@
 clusters. As a result, ISODS can be used to generate clusters for other
-clustering algorithms, such as K-Means. In the next section, ISODS is applied to
-both ISODATA and K-Means on some benchmark UCI datasets [CITE] and compared with
-Random Centers and K-Means++ Centers.
+clustering algorithms, such as K-Means.
 
@@
 summarized with the mean and standard deviation of that class, as well as the
-number of members in each class. To determine $\theta n$, the minimum number of
-cluster members used for discarding, we find the class with the fewest members
-and use the number of members in that class as $\theta n$. For $\theta e$, the
-maximum standard deviation allowed for a cluster, we follow a similar approach
-and find the maximum standard deviation within a class. Finally, for $\theta c$,
-we calculate the distance between the means of each class, and use the minimum
-distance as $\theta c$.
+number of members in each class. Typically, since we are working with
+unsupervised learning algorithms, class information may not be available.
+
+To determine $\theta n$, the minimum number of cluster members used for
+discarding, we find the class with the fewest members and use the number of
+members in that class as $\theta n$. For $\theta e$, the maximum standard
+deviation allowed for a cluster, we follow a similar approach and find the
+maximum standard deviation within a class. Finally, for $\theta c$, we calculate
+the distance between the means of each class, and use the minimum distance as
+$\theta c$.
+
+For experiments 1 and 2, we explore applying ISODS to ISODATA and K-Means by
+using benchmark UCI datasets Iris, Wine, and Glass. We also show results from
+applying Random Centers and K-Means++ Centers. To evaluate the cluster results,
+external cluster validation measures are used. We utilize pair-counting measures
+which rely on counting matching pairs between two sets of data to determine
+agreement and disagreement \cite{Rezaei2016}. The pair-counting measures applied
+are \textit{rand index} (RI - agreement probability), \textit{adjusted rand
+  index} (ARI), \textit{mirkin's index} (MI - disagreement probability), and
+\textit{hubert's index} (HI - $P(agree) - P(disagree)$)\cite{Hubert1985}
+\cite{Meila2003}.
 
@@
 
-A statement requiring citation \cite{Figueredo:2009dg}.
 \blindtext % Dummy text
@@
 
-\begin{thebibliography}{99} % Bibliography - this is intentionally simple in this template
-
-\bibitem[Figueredo and Wolf, 2009]{Figueredo:2009dg}
-Figueredo, A.~J. and Wolf, P. S.~A. (2009).
-\newblock Assortative pairing and life history strategy - a cross-cultural
-  study.
-\newblock {\em Human Nature}, 20:317--330.
-
-\end{thebibliography}
+\bibliographystyle{unsrt}
+\bibliography{paper}
 

### 2021-01-08 paper/paper.tex
modify style to match conference requirements

@@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Journal Article
-% LaTeX Template
-% Version 1.4 (15/5/16)
-%
-% This template has been downloaded from:
-% http://www.LaTeXTemplates.com
-%
-% Original author:
-% Frits Wenneker (http://www.howtotex.com) with extensive modifications by
-% Vel (vel@LaTeXTemplates.com)
-%
-% License:
-% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\documentclass[a4paper,10pt]{article}
+\usepackage{graphicx}
+\begin{document}
 
-%----------------------------------------------------------------------------------------
-%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
-%----------------------------------------------------------------------------------------
+\title{\bf Patent Duration, Innovative Performance, and Technology Diffusion}
+\author{Noah Gardner$^1$,
+      Sahar Yarmohammadtoosky$^2$, \\
+      Anthony Phan$^3$,
+      and Chih-Cheng Hung$^4$\\[1ex]
+      \sl \small Kennesaw State University,
+      Marietta, GA, USA$^{1, 3, 4}$\\
+      \sl \small Iran University of Science and Technology,
+      Tehran, Iran$^2$\\[1ex]
+      \small $^1$ngardn10@students.kennesaw.edu, 
+      \small $^2$yarmohamadishr@gmail.com, \\
+      \small $^3$aphan5@students.kennesaw.edu,
+      \small $^4$chung1@kennesaw.edu, \\}
+\date{}
+\maketitle
 
-\documentclass[twoside,twocolumn]{article}
-
-\usepackage{blindtext} % Package to generate dummy text throughout this template
-
-\usepackage[sc]{mathpazo} % Use the Palatino font
-\usepackage[T1]{fontenc} % Use 8-bit encoding that has 256 glyphs
-\linespread{1.05} % Line spacing - Palatino needs more space between lines
-\usepackage{microtype} % Slightly tweak font spacing for aesthetics
-
-\usepackage[english]{babel} % Language hyphenation and typographical rules
-
-\usepackage[hmarginratio=1:1,top=32mm,columnsep=20pt]{geometry} % Document margins
-\usepackage[hang, small,labelfont=bf,up,textfont=it,up]{caption} % Custom captions under/above floats in tables or figures
-\usepackage{booktabs} % Horizontal rules in tables
-
-\usepackage{lettrine} % The lettrine is the first enlarged letter at the beginning of the text
-
-\usepackage{enumitem} % Customized lists
-\setlist[itemize]{noitemsep} % Make itemize lists more compact
-
-\usepackage{abstract} % Allows abstract customization
-\renewcommand{\abstractnamefont}{\normalfont\bfseries} % Set the "Abstract" text to bold
-\renewcommand{\abstracttextfont}{\normalfont\small} % Set the abstract itself to small italic text
-
-\usepackage{titlesec} % Allows customization of titles
-\renewcommand\thesection{\Roman{section}} % Roman numerals for the sections
-\renewcommand\thesubsection{\Roman{subsection}} % roman numerals for subsections
-\titleformat{\section}[block]{\large\scshape\centering}{\thesection.}{1em}{} % Change the look of the section titles
-\titleformat{\subsection}[block]{\large}{\thesubsection.}{1em}{} % Change the look of the section titles
-
-\usepackage{fancyhdr} % Headers and footers
-\pagestyle{fancy} % All pages have headers and footers
-\fancyhead{} % Blank out the default header
-\fancyfoot{} % Blank out the default footer
-\fancyhead[C]{Running title $\bullet$ May 2016 $\bullet$ Vol. XXI, No. 1} % Custom header text
-\fancyfoot[RO,LE]{\thepage} % Custom footer text
-
-\usepackage{titling} % Customizing the title section
-
-\usepackage{hyperref} % For hyperlinks in the PDF
-
-%----------------------------------------------------------------------------------------
-%	TITLE SECTION
-%----------------------------------------------------------------------------------------
-
-\setlength{\droptitle}{-4\baselineskip} % Move the title up
-
-\pretitle{\begin{center}\Huge\bfseries} % Article title formatting
-\posttitle{\end{center}} % Article title closing formatting
-\title{ISODS: An ISODATA-Based Initial Centroid Algorithm} % Article title
-\author{%
-  \textsc{
-    Noah Gardner$^1$\thanks{ngardn10@students.kennesaw.edu}, 
-    Sahar Yarmohammadtoosky$^2$\thanks{yarmohamadishr@gmail.com},
-    }\\[1ex]
-  \textsc{
-    Anthony Phan$^1$\thanks{aphan5@students.kennesaw.edu},
-    Chih-Cheng Hung$^1$\thanks{chung1@students.kennesaw.edu}} \\[1ex]
-  \normalsize $^1$Kennesaw State University \\
-  \normalsize College of Computing and Software Engineering \\
-  \normalsize Marietta, GA, USA \\[1ex]
-  \normalsize $^2$Iran University of Science and Technology \\
-  \normalsize Tehran, Iran
-}
-\date{} % Leave empty to omit a date
-
-\renewcommand{\maketitlehookd}{%
 \begin{abstract}
@@
       Iterative Self-Organizing Data Analysis Technique (ISODATA) is a clustering
-algorithm based on K-Means that intends to find the correct number of clusters
-through merging and splitting. Compared to K-Means, the ISODATA algorithm should
-be resilient to initial cluster centers. However, ISODATA is also sensitive to
-initial cluster centers which impacts the performance of the algorithm. This
-work proposes using the ISODATA approach of splitting clusters to select quality
-initial cluster centers. The proposed approach is evaluated on several benchmark
-UCI datasets as well as the Salinas-A hyperspectral image dataset. The results
-are evaluated using standard clustering evaluation parameters. \\ \\
-\textit{\textbf{Keywords:} Clustering, ISODATA, Initial Centroid, Remote Sensing} \\
+      algorithm based on K-Means that intends to find the correct number of
+      clusters through merging and splitting. Compared to K-Means, the ISODATA
+      algorithm should be resilient to initial cluster centers. However, ISODATA
+      is also sensitive to initial cluster centers which impacts the performance
+      of the algorithm. This work proposes using the ISODATA approach of splitting
+      clusters to select quality initial cluster centers. The proposed approach is
+      evaluated on several benchmark UCI datasets as well as the Salinas-A
+      hyperspectral image dataset. The results are evaluated using standard
+      clustering evaluation parameters. \\
+      {\bf Keywords :} Clustering, ISODATA, Initial Centroid, Remote Sensing
 \end{abstract}
-}
 
-%----------------------------------------------------------------------------------------
-
-\begin{document}
-
-% Print the title
-\maketitle
-
-%----------------------------------------------------------------------------------------
-%	ARTICLE CONTENTS
-%----------------------------------------------------------------------------------------
+% =========================================================================== %
 
@@
 
-\lettrine[nindent=0em,lines=3]{R} emote sensing technology is used when the data
-collection takes place at a remote distance from the subject matter, such as the
-study of daily weather and long-term climate change, land cover monitoring,
-ecosystem modeling of vegetation, and so on \cite{Jensen1996,Mallapragada2018}.
-Multispectral imaging technology is one of the instruments for remote sensing
-used since the late 1960s \cite{Khan2018,Landgrebe1999}. However, multispectral
-imaging systems use a low number of spectral bands (three to six spectral bands
-in different systems). Increasing spectral resolution in the hyperspectral
-images (HSI) with several hundred spectral bands of the observational scene in a
-single acquisition makes it possible to acquire detailed examination of
-observational scenes \cite{Khan2018}. The ground-based version of hyperspectral
-images has gained immense interest. It has been applied to numerous electronic
-imaging applications, including food inspection, forensic science, medical
-surgery and diagnosis, and military applications
+Remote sensing technology is used when the data collection takes place at a
+remote distance from the subject matter, such as the study of daily weather and
+long-term climate change, land cover monitoring, ecosystem modeling of
+vegetation, and so on \cite{Jensen1996,Mallapragada2018}. Multispectral imaging
+technology is one of the instruments for remote sensing used since the late
+1960s \cite{Khan2018,Landgrebe1999}. However, multispectral imaging systems use
+a low number of spectral bands (three to six spectral bands in different
+systems). Increasing spectral resolution in the hyperspectral images (HSI) with
+several hundred spectral bands of the observational scene in a single
+acquisition makes it possible to acquire detailed examination of observational
+scenes \cite{Khan2018}. The ground-based version of hyperspectral images has
+gained immense interest. It has been applied to numerous electronic imaging
+applications, including food inspection, forensic science, medical surgery and
+diagnosis, and military applications
 \cite{Govender2009,Adam2009,Fischer2006,Carrasco2003}. Because of the difficulty
@@
 
-% The paper is organized as follows: Section 2 introduces K-Means and ISODATA
-% clustering algorithms. The new proposed algorithm is explained in Section 3. We
-% present the experimental results in Section 4. Finally, Section 5 contains
-% concluding remarks.
+The paper is organized as follows: Section 2 introduces K-Means and ISODATA
+clustering algorithms. The new proposed algorithm is explained in Section 3. We
+present the experimental results in Section 4. Finally, Section 5 contains
+concluding remarks.
 
-%------------------------------------------------
+% =========================================================================== %
+
+\section{Background}
 
@@
       \item \textbf{Step 1:} Choose $K$ cluster centers from the data set.
-  \item \textbf{Step 2:} Assign each element of the data set to the cluster based
-        on the minimum distance from the cluster center.
-  \item \textbf{Step 3:} Re-calculate new cluster centers for each cluster after
-        the assignment in \textbf{Step 2} by computing the mean of all patterns
-        assigned to each cluster.
+      \item \textbf{Step 2:} Assign each element of the data set to the cluster
+            based on the minimum distance from the cluster center.
+      \item \textbf{Step 3:} Re-calculate new cluster centers for each cluster
+            after the assignment in \textbf{Step 2} by computing the mean of all
+            patterns assigned to each cluster.
       \item \textbf{Step 4:} Repeat \textbf{Step 2 and Step 3} until clustering
@@
 initial values can result in a different classification.
-
-%------------------------------------------------
 
@@
 \begin{itemize}
-  \item $\theta n$ - minimum number of cluster members, which determines whether
-        the cluster is discarded,
+      \item $\theta n$ - minimum number of cluster members, which determines
+            whether the cluster is discarded,
       \item $\theta e$ - maximum standard deviation allowed for a cluster, which
             determines whether the cluster is split,
-  \item $\theta c$ - minimum distance between clusters, which determines whether
-        two clusters are merged.
+      \item $\theta c$ - minimum distance between clusters, which determines
+            whether two clusters are merged.
 \end{itemize}
@@
       \item \textbf{Step 1:} Choose $K$ cluster centers from the data set.
-  \item \textbf{Step 2:} Assign each element of the dataset to the cluster based
-        on the minimum distance from the cluster center.
+      \item \textbf{Step 2:} Assign each element of the dataset to the cluster
+            based on the minimum distance from the cluster center.
       \item \textbf{Step 3:} Re-calculate new cluster centers for each cluster by
@@
       \item \textbf{Step 5:} If this iteration is an odd numbered iteration or the
-        number of clusters is less than or equal to  $1/2 * K$, then split clusters
-        which have a standard deviation in a dimension greater than ($\theta e$). Move
-        to \textbf{Step 7}.
-  \item \textbf{Step 6:} If this iteration is an even numbered iteration or the
-        number of clusters is greater than or equal to $2 * K$, then merge clusters which
-        have an intercluster distance less than ($\theta c$).
+            number of clusters is less than or equal to  $1/2 * K$, then split
+            clusters which have a standard deviation in a dimension greater than
+            ($\theta e$). Move to \textbf{Step 7}.
+      \item \textbf{Step 6:} If this iteration is an even numbered iteration or
+            the number of clusters is greater than or equal to $2 * K$, then merge
+            clusters which have an intercluster distance less than ($\theta c$).
       \item \textbf{Step 7:} If this is not the last iteration, start the process
-        again at step 2. If this is the last iteration or clustering convergence occurs,
-        then end the process.
+            again at step 2. If this is the last iteration or clustering
+            convergence occurs, then end the process.
 \end{itemize}
@@
 
-%------------------------------------------------
+% =========================================================================== %
 
@@
 \begin{itemize}
-  \item \textbf{Step 1:} Choose $1$ observation as a cluster center from the dataset.
-  \item \textbf{Step 2:} Assign each element of the dataset to the cluster based
-        on the minimum distance from the cluster center.
+      \item \textbf{Step 1:} Choose $1$ observation as a cluster center from the
+            dataset.
+      \item \textbf{Step 2:} Assign each element of the dataset to the cluster
+            based on the minimum distance from the cluster center.
       \item \textbf{Step 3:} Re-calculate new cluster centers for each cluster by
@@
       \item \textbf{Step 4:} If the number of clusters is equal to $K$, then the
-        process is terminated. If the number of clusters is greater than $K$, then the
-        clusters with the fewest members is discarded, and the algorithm moves to
-        \textbf{Step 2}.
-  \item \textbf{Step 5:} Calculate the standard deviation of each cluster in each
-        dimension. Then, split each cluster along the dimension with the greatest
-        standard deviation.
+            process is terminated. If the number of clusters is greater than $K$,
+            then the clusters with the fewest members is discarded, and the
+            algorithm moves to \textbf{Step 2}.
+      \item \textbf{Step 5:} Calculate the standard deviation of each cluster in
+            each dimension. Then, split each cluster along the dimension with the
+            greatest standard deviation.
 \end{itemize}
@@
 
-%------------------------------------------------
+% =========================================================================== %
+
 \section{Results}
@@
       index} (ARI), \textit{mirkin's index} (MI - disagreement probability), and
-\textit{hubert's index} (HI - $P(agree) - P(disagree)$)\cite{Hubert1985}
-\cite{Meila2003}.
+\textit{hubert's index} (HI - $P(agree) - P(disagree)$)\cite{Hubert1985,
+      Meila2003}.
 
-%------------------------------------------------
+% =========================================================================== %
 
 \section{Conclusion}
+Put Conclusion here.
 
-\subsection{Subsection One}
+% =========================================================================== %
 
-\blindtext % Dummy text
+\subsection*{Acknowledgements}
+Put Acknowledgements here.
 
-\subsection{Subsection Two}
+\vspace{5mm}
 
-\blindtext % Dummy text
-
-%----------------------------------------------------------------------------------------
-%	REFERENCE LIST
-%----------------------------------------------------------------------------------------
+\begin{center}
+      {\large \bf References}
+\end{center}
 
@@
 
-%----------------------------------------------------------------------------------------
-
 \end{document}

### 2021-01-08 paper/paper.tex
missed title update
experiment 1 table and results

@@
 
-\title{\bf Patent Duration, Innovative Performance, and Technology Diffusion}
-\author{Noah Gardner$^1$,
-      Sahar Yarmohammadtoosky$^2$, \\
-      Anthony Phan$^3$,
-      and Chih-Cheng Hung$^4$\\[1ex]
-      \sl \small Kennesaw State University,
-      Marietta, GA, USA$^{1, 3, 4}$\\
-      \sl \small Iran University of Science and Technology,
-      Tehran, Iran$^2$\\[1ex]
-      \small $^1$ngardn10@students.kennesaw.edu, 
-      \small $^2$yarmohamadishr@gmail.com, \\
-      \small $^3$aphan5@students.kennesaw.edu,
-      \small $^4$chung1@kennesaw.edu, \\}
+\title{\bf ISODS: An ISODATA-Based Initial Centroid Algorithm}
+\author{Noah Gardner$^1$, Sahar Yarmohammadtoosky$^2$, \\
+      Anthony Phan$^3$, and Chih-Cheng Hung$^4$\\[1ex]
+      \sl \small Kennesaw State University, Marietta, GA, USA$^{1, 3, 4}$\\
+      \sl \small Iran University of Science and Technology, Tehran,
+      Iran$^2$\\[1ex]
+      \small $^1$ngardn10@students.kennesaw.edu, \small
+      $^2$yarmohamadishr@gmail.com, \\
+      \small $^3$aphan5@students.kennesaw.edu, \small $^4$chung1@kennesaw.edu,
+      \\}
 \date{}
@@
       \noindent 
-      Iterative Self-Organizing Data Analysis Technique (ISODATA) is a clustering
-      algorithm based on K-Means that intends to find the correct number of
-      clusters through merging and splitting. Compared to K-Means, the ISODATA
-      algorithm should be resilient to initial cluster centers. However, ISODATA
-      is also sensitive to initial cluster centers which impacts the performance
-      of the algorithm. This work proposes using the ISODATA approach of splitting
-      clusters to select quality initial cluster centers. The proposed approach is
-      evaluated on several benchmark UCI datasets as well as the Salinas-A
-      hyperspectral image dataset. The results are evaluated using standard
-      clustering evaluation parameters. \\
-      {\bf Keywords :} Clustering, ISODATA, Initial Centroid, Remote Sensing
+      Iterative Self-Organizing Data Analysis Technique (ISODATA) is a
+      clustering algorithm based on K-Means that intends to find the correct
+      number of clusters through merging and splitting. Compared to K-Means, the
+      ISODATA algorithm should be resilient to initial cluster centers. However,
+      ISODATA is also sensitive to initial cluster centers which impacts the
+      performance of the algorithm. This work proposes using the ISODATA
+      approach of splitting clusters to select quality initial cluster centers.
+      The proposed approach is evaluated on several benchmark UCI datasets as
+      well as the Salinas-A hyperspectral image dataset. The results are
+      evaluated using standard clustering evaluation parameters. \\
+      {\bf Keywords :} Clustering, ISODATA, Initial Centroids, Remote Sensing
 \end{abstract}
@@
             based on the minimum distance from the cluster center.
-      \item \textbf{Step 3:} Re-calculate new cluster centers for each cluster by
-            computing the mean of all patterns assigned to each cluster.
+      \item \textbf{Step 3:} Re-calculate new cluster centers for each cluster
+            by computing the mean of all patterns assigned to each cluster.
       \item \textbf{Step 4:} Discard clusters that have fewer than ($\theta n$)
             members. If any clusters were discarded, move to \textbf{Step 7}.
-      \item \textbf{Step 5:} If this iteration is an odd numbered iteration or the
-            number of clusters is less than or equal to  $1/2 * K$, then split
-            clusters which have a standard deviation in a dimension greater than
-            ($\theta e$). Move to \textbf{Step 7}.
+      \item \textbf{Step 5:} If this iteration is an odd numbered iteration or
+            the number of clusters is less than or equal to  $1/2 * K$, then
+            split clusters which have a standard deviation in a dimension
+            greater than ($\theta e$). Move to \textbf{Step 7}.
       \item \textbf{Step 6:} If this iteration is an even numbered iteration or
-            the number of clusters is greater than or equal to $2 * K$, then merge
-            clusters which have an intercluster distance less than ($\theta c$).
-      \item \textbf{Step 7:} If this is not the last iteration, start the process
-            again at step 2. If this is the last iteration or clustering
+            the number of clusters is greater than or equal to $2 * K$, then
+            merge clusters which have an intercluster distance less than
+            ($\theta c$).
+      \item \textbf{Step 7:} If this is not the last iteration, start the
+            process again at step 2. If this is the last iteration or clustering
             convergence occurs, then end the process.
@@
 
-There are some disadvantages to the ISODATA algorithm. First, ISODATA is
-sensitive to the parameters $\theta n$, $\theta e$, and $\theta c$. These
-parameters should be fine-tuned for optimal performance based on the attributes
-of the dataset being tested. In addition, similar to K-Means, it is sensitive to
-initial cluster centers, although less so than K-Means due to the extra
-processes of discarding, splitting, and merging. The initial centroids can
-affect both the quality of clusters produced and the number of iterations
-required to converge [CITE].
+There are some disadvantages to the ISODATA algorithm. First, it is sensitive to
+the parameters $\theta n$, $\theta e$, and $\theta c$. These parameters should
+be fine-tuned for optimal performance based on the attributes of the dataset
+being tested. In addition, similar to K-Means, it is sensitive to initial
+cluster centers, although less so than K-Means due to the extra processes of
+discarding, splitting, and merging. The initial centroids can affect both the
+quality of clusters produced and the number of iterations required to converge
+      [CITE].
 
@@
 The proposed clustering algorithm's objective is to improve the result we obtain
-from the ISODATA algorithm by targeting the initial cluster centers. Inspired by
-the ISODATA algorithm, Iterative Self-Organizing Data Split (ISODS) uses the
-splitting process from ISODATA to create the required initial cluster centers to
-use for the ISODATA algorithm. The inputs for the algorithm are the number of
-cluster centers $K$ as an integer, and the data to generate centers from. The
-algorithm starts by selecting $1$ sample as an initial cluster center. In the
-next step, each data is assigned to the cluster and the mean of the cluster is
-updated. Then, by the ISODATA splitting process, each cluster is split along the
-dimension with the highest standard deviation. The assignments for the new
+from the clustering algorithms by targeting the initial cluster centers.
+Inspired by the ISODATA algorithm, Iterative Self-Organizing Data Split (ISODS)
+uses the splitting process from ISODATA to create the required initial cluster
+centers to use for a clustering algorithm. The inputs for the algorithm are the
+number of cluster centers $K$ as an integer, and the data to generate centers
+from.
+
+The algorithm starts by selecting $1$ sample as an initial cluster center. In
+the next step, each data is assigned to the cluster and the mean of the cluster
+is updated. Then, by the ISODATA splitting process, each cluster is split along
+the dimension with the highest standard deviation. The assignments for the new
 clusters are calculated based on distance and the centers are updated. This
@@
             based on the minimum distance from the cluster center.
-      \item \textbf{Step 3:} Re-calculate new cluster centers for each cluster by
-            computing the mean of all patterns assigned to each cluster.
+      \item \textbf{Step 3:} Re-calculate new cluster centers for each cluster
+            by computing the mean of all patterns assigned to each cluster.
       \item \textbf{Step 4:} If the number of clusters is equal to $K$, then the
-            process is terminated. If the number of clusters is greater than $K$,
-            then the clusters with the fewest members is discarded, and the
+            process is terminated. If the number of clusters is greater than
+            $K$, then the clusters with the fewest members is discarded, and the
             algorithm moves to \textbf{Step 2}.
       \item \textbf{Step 5:} Calculate the standard deviation of each cluster in
-            each dimension. Then, split each cluster along the dimension with the
-            greatest standard deviation.
+            each dimension. Then, split each cluster along the dimension with
+            the greatest standard deviation.
 \end{itemize}
@@
 \section{Results}
+
+\begin{table}
+      \begin{tabular}{ |p{1.25cm}||p{1.25cm} p{1.25cm} p{1.25cm} p{1.25cm} p{1.25cm} p{1.25cm}|}
+            \hline
+            \multicolumn{7}{|c|}{\textbf{ISODATA}}                               \\
+            \hline
+            Init. & Iter.  & Clusters & ARI             & RI    & MI    & HI     \\
+            \hline
+            \multicolumn{7}{|c|}{\textit{Iris}}                                  \\[1ex]
+            ISODS & 13.46  & 2.599    & \textbf{0.646}  & 0.826 & 0.173 & 0.653  \\
+            KM++  & 15.43  & 2.544    & 0.628           & 0.816 & 0.184 & 0.631  \\
+            Rand  & 11.90  & 2.41     & 0.607           & 0.802 & 0.198 & 0.604  \\
+            \hline
+            \multicolumn{7}{|c|}{\textit{Glass (Normalized)}}                    \\[1ex]
+            ISODS & 3.19   & 1.066    & \textbf{0.0018} & 0.266 & 0.734 & -0.468 \\
+            KM++  & 4.02   & 1.028    & 0.0011          & 0.265 & 0.735 & -0.470 \\
+            Rand  & 3.69   & 1.055    & 0.0013          & 0.266 & 0.733 & -0.467 \\
+            \hline
+            \multicolumn{7}{|c|}{\textit{Wine}}                                  \\[1ex]
+            ISODS & 9.16   & 1.836    & \textbf{0.313}  & 0.618 & 0.382 & 0.236  \\
+            KM++  & 8.76   & 1.816    & 0.307           & 0.612 & 0.388 & 0.224  \\
+            Rand  & 9.26   & 1.810    & 0.304           & 0.608 & 0.391 & 0.217  \\
+            \hline 
+            \multicolumn{7}{|c|}{\textbf{K-Means}}                               \\
+            \hline
+            \multicolumn{7}{|c|}{\textit{Iris}}                                  \\[1ex]
+            ISODS & 2.29   & 3        & \textbf{0.623}  & 0.833 & 0.167 & 0.667  \\
+            KM++  & 4.149  & 3        & 0.584           & 0.811 & 0.189 & 0.621  \\
+            Rand  & 29.402 & 3        & 0.569           & 0.797 & 0.203 & 0.593  \\
+            \hline
+            \multicolumn{7}{|c|}{\textit{Glass (Normalized)}}                    \\[1ex]
+            ISODS & 5.651  & 6        & 0.164           & 0.665 & 0.335 & 0.329  \\
+            KM++  & 4.866  & 6        & \textbf{0.173}  & 0.643 & 0.357 & 0.285  \\
+            Rand  & 77.178 & 6        & 0.151           & 0.634 & 0.366 & 0.268  \\
+            \hline
+            \multicolumn{7}{|c|}{\textit{Wine}}                                  \\[1ex]
+            ISODS & 3.145  & 3        & \textbf{0.882}  & 0.947 & 0.053 & 0.895  \\
+            KM++  & 5.388  & 3        & 0.839           & 0.926 & 0.074 & 0.851  \\
+            Rand  & 24.775 & 3        & 0.834           & 0.920 & 0.080 & 0.840  \\
+            \hline
+      \end{tabular}
+      \caption{UCI benchmark dataset results averaged over 1000 iterations of
+            ISODATA and K-Means comparing three different initial cluster center
+            algorithms (\textit{Init.}). Results in the table contain
+            information on average iterations required (\textit{Iter.}), average
+            number of clusters created (\textit{Clusters}), and validity
+            measures.}
+      \label{tab:ucitable}
+\end{table}
 
@@
 
-For experiments 1 and 2, we explore applying ISODS to ISODATA and K-Means by
-using benchmark UCI datasets Iris, Wine, and Glass. We also show results from
-applying Random Centers and K-Means++ Centers. To evaluate the cluster results,
-external cluster validation measures are used. We utilize pair-counting measures
-which rely on counting matching pairs between two sets of data to determine
-agreement and disagreement \cite{Rezaei2016}. The pair-counting measures applied
-are \textit{rand index} (RI - agreement probability), \textit{adjusted rand
-      index} (ARI), \textit{mirkin's index} (MI - disagreement probability), and
-\textit{hubert's index} (HI - $P(agree) - P(disagree)$)\cite{Hubert1985,
-      Meila2003}.
+For the first experimental results shown in Table \ref{tab:ucitable}, we explore
+applying ISODS to ISODATA and K-Means by using benchmark UCI \textit{Iris,}
+\textit{Wine,} and \textit{Glass} datasets \cite{Dua2019}. We also show results
+from applying Random Centers and K-Means++ Centers. For the \textit{Glass}
+dataset, the data was Z-Score normalized (based on Z-Score of the training data)
+because ISODATA can have problems converging on the raw dataset. We used $80\%$
+of the data for training and $20\%$ of the data for testing to evaluate the
+models. Following standard practice, the models were trained on the training
+data, then fit to the test data. For the ISODATA parameters, we used the
+approach of summarizing the class data based on the information in the training
+data. The output labels (cluster assignments) were compared to the ground truth
+labels contained in the test data using pair-counting measures.
+
+To evaluate the cluster results, external cluster validation measures are used.
+We utilize pair-counting measures which rely on counting matching pairs between
+two sets of data to determine agreement and disagreement \cite{Rezaei2016}. The
+pair-counting measures applied are \textit{rand index} (RI - agreement
+probability), \textit{adjusted rand index} (ARI), \textit{mirkin's index} (MI -
+disagreement probability), and \textit{hubert's index} (HI - $(P(agree)$ -
+$P(disagree))$) \cite{Hubert1985, Meila2003}. When comparing ground truth labels
+to cluster assignments, RI can serve as an estimation of the accuracy. ARI is
+similar to the RI, though it is corrected for chance and thus does not count
+randomly generated labels \cite{Rand1971}.
 
@@
 \subsection*{Acknowledgements}
-Put Acknowledgements here.
+This research was supported by the College of Computing and Software Engineering
+at Kennesaw State University. We also want to thank our Laboratory for Machine
+Vision and Security Research for providing us resources and mentoring.
 

### 2021-01-08 paper/paper.tex
add experiment 2 results and conclusion

@@
 \usepackage{graphicx}
+\usepackage{caption}
+\usepackage{subcaption}
+
+\graphicspath{ {./img/} }
 \begin{document}
@@
 
-\section{Results}
-
-\begin{table}
+\begin{table}[ht]
       \begin{tabular}{ |p{1.25cm}||p{1.25cm} p{1.25cm} p{1.25cm} p{1.25cm} p{1.25cm} p{1.25cm}|}
@@
 
+\section{Results}
+
 Due to the response of the ISODATA algorithm to the discarding, merging, and
@@
 
+\begin{figure}[ht]
+      \centering
+      \begin{subfigure}[b]{0.4\textwidth}
+            \centering
+            \includegraphics[width=\linewidth]{salinasgt.png}
+            \caption{Salinas-A Ground Truth}
+            \label{fig:salinasgt}
+      \end{subfigure}
+      \begin{subfigure}[b]{0.4\textwidth}
+            \centering
+            \includegraphics[width=\linewidth]{risodata.png}
+            \caption{Random Centers (ARI=0.797)}
+            \label{fig:salinasr}
+      \end{subfigure}\hfill
+      \centering
+      \begin{subfigure}[b]{0.4\textwidth}
+            \centering
+            \includegraphics[width=\linewidth]{kisodata.png}
+            \caption{KM++ Centers (ARI=0.781)}
+            \label{fig:salinask}
+      \end{subfigure}
+      \begin{subfigure}[b]{0.4\textwidth}
+            \centering
+            \includegraphics[width=\linewidth]{iisodata.png}
+            \caption{ISODS Centers (ARI=0.758)}
+            \label{fig:salinasi}
+      \end{subfigure}\hfill \caption{Heatmap showing cluster assignments of each
+            pixel using ISODATA with different initial cluster center methods.}
+      \label{fig:salinas}
+\end{figure}
+
+Following the results on the benchmark datasets, we applied ISODATA to the
+Salinas-A hyperspectral image dataset \cite{Salinas}. Salinas-A is a small
+subscene of the Salinas image. It has a size of $86$ pixels by $83$ pixels, each
+pixel contains $224$ bands, and each pixel is assigned to $1$ of $6$ classes,
+disregarding the background. Due to the large number of dimensions in this
+dataset, prinicipal component analysis (PCA) is used to extract features
+\cite{Rodarmel2002}. In this experimental setup, we use $3$ prinicipal
+components as the input to the model. We utilize the same 80/20 test-train split
+as the previous experiment and Z-Score Normalize the data. Additionally, the
+same ISODATA parameter selection approach is employed. Then, we run ISODATA for
+$1000$ iterations using ISODS, K-Means++ Centers, and Random Centers. We save
+the best performing model (based on test data performance) for each initial
+cluster center approach and fit each pixel to the original dataset to obtain the
+pictures shown in Figure \ref{fig:salinas}. Finally, the external cluster
+validation measures are shown in Table \ref{tab:salinastable}.
+
+\begin{table}[ht]
+      \begin{tabular}{ |p{1.25cm}||p{1.25cm} p{1.25cm} p{1.25cm} p{1.25cm} p{1.25cm} p{1.25cm}|}
+            \hline
+            \multicolumn{7}{|c|}{\textbf{ISODATA} - \textit{Salinas-A}}        \\
+            \hline
+            Init. & Iter. & Clusters & ARI            & RI    & MI    & HI     \\
+            \hline
+            \multicolumn{7}{|c|}{\textit{Average}}                             \\[1ex]
+            ISODS & 2.772 & 2.42     & \textbf{0.156} & 0.388 & 0.612 & -0.224 \\
+            KM++  & 3.177 & 1.097    & 0.019          & 0.228 & 0.772 & -0.544 \\
+            Rand  & 3.693 & 1.584    & 0.084          & 0.308 & 0.691 & -0.383 \\
+            \hline
+            \multicolumn{7}{|c|}{\textit{Best}}                                \\[1ex]
+            ISODS & 2     & 6        & 0.758          & 0.921 & 0.078 & 0.842  \\
+            KM++  & 8     & 6        & 0.781          & 0.930 & 0.070 & 0.860  \\
+            Rand  & 14    & 6        & \textbf{0.798} & 0.933 & 0.066 & 0.867  \\
+            \hline
+      \end{tabular}
+      \caption{Salinas-A dataset average and best results over 1000 iterations
+            of ISODATA comparing three different initial cluster center
+            algorithms (\textit{Init.}). Results in the table contain
+            information on iterations required (\textit{Iter.}), number of
+            clusters created (\textit{Clusters}), and validity measures.}
+      \label{tab:salinastable}
+\end{table}
+
 % =========================================================================== %
@@
 \section{Conclusion}
-Put Conclusion here.
+This study presents a method for generating initial cluster centers based on a
+splitting approach, ISODS. We tested ISODS on bechmark datasets as well as the
+Salinas-A hyperspectral image dataset. Both tests on average shows the
+performance of ISODS using external cluster validation measures, especially on
+the \textit{adjusted rand index} measure. Additonally, ISODS may decrease the
+number of iterations required for a clustering algorithm such as ISODATA or
+K-Means to converge. However, due to the complexity of the ISODS approach,
+specifically calculating the means of the clusters, the number of iterations
+taken by ISODS should be considered. When used in conjunction with ISODATA,
+ISODS also shows an average number of clusters generated closer to the expected
+output.
 
@@
 
-\begin{center}
-      {\large \bf References}
-\end{center}
-
 \bibliographystyle{unsrt}

### 2021-01-09 paper/paper.tex
paper revision 1

@@
-\documentclass[a4paper,10pt]{article}
+\documentclass[a4paper, 10pt]{article}
 \usepackage{graphicx}
@@
       Iterative Self-Organizing Data Analysis Technique (ISODATA) is a
-      clustering algorithm based on K-Means that intends to find the correct
-      number of clusters through merging and splitting. Compared to K-Means, the
-      ISODATA algorithm should be resilient to initial cluster centers. However,
-      ISODATA is also sensitive to initial cluster centers which impacts the
-      performance of the algorithm. This work proposes using the ISODATA
-      approach of splitting clusters to select quality initial cluster centers.
-      The proposed approach is evaluated on several benchmark UCI datasets as
-      well as the Salinas-A hyperspectral image dataset. The results are
-      evaluated using standard clustering evaluation parameters. \\
+      clustering algorithm based on the K-Means algorithm that intends to find
+      the correct number of clusters through merging and splitting. It is
+      difficult to determine the correct number of clusters in many real-world
+      applications, in particular, remote sensing image interpretation. Compared
+      to K-Means, the ISODATA algorithm should be resilient to initial cluster
+      centers. However, ISODATA is also sensitive to initial cluster centers
+      which impacts the performance of the algorithm. This work proposes using
+      the ISODATA approach of splitting clusters to select quality initial
+      cluster centers, called Iterative Self-Organizing Data Split (ISODS). The
+      proposed approach is evaluated on several benchmark UCI datasets as well
+      as the Salinas-A hyperspectral image dataset. The results are evaluated
+      using standard clustering evaluation parameters. When ISODS is used
+      together with ISODATA, the performance is improved and the generated
+      clusters more closely resemble the expected number of clusters.\\
       {\bf Keywords :} Clustering, ISODATA, Initial Centroids, Remote Sensing
@@
 understand the data set better. There are different algorithms for Clustering,
-such as K-Means, fuzzy C-Means, ISODATA \cite{Ball1965}, and so on. In the
-K-Means algorithm, the goal is to minimize the sum of the squared distances
-points have in the data set from its center. The problem with the K-Means
-algorithm is that it requires a priori knowledge about the number of classes in
-the data set, and choosing the cluster centers randomly may not result in
-optimal partitioning of the data set.
+such as K-Means \cite{Lloyd1982,Macqueen1967}, fuzzy C-Means, ISODATA
+\cite{Ball1965}, and so on. In the K-Means algorithm, the goal is to minimize
+the sum of the squared distances points have in the data set from its center.
+The problem with the K-Means algorithm is that it requires a priori knowledge
+about the number of classes in the data set, and choosing the cluster centers
+randomly may not result in optimal partitioning of the data set.
 
@@
 The K-Means algorithm is an unsupervised clustering algorithm and aims to
-partition $N$ observations into $K$ clusters [CITE]. Data points in the same
-cluster are very similar to each other and have less variation compared to data
-points assigned to other clusters. The user defines $K$, the number of expected
-clusters. Following is a description of the K-Means algorithm:
+partition $N$ observations into $K$ clusters \cite{Lloyd1982,Macqueen1967}. Data
+points in the same cluster are very similar to each other and have less
+variation compared to data points assigned to other clusters. The user defines
+$K$, the number of expected clusters. Following is a description of the K-Means
+algorithm:
 
@@
 For selecting the initial cluster centers in \textbf{Step 1}, there are two
-standard methods: Random Centers, and K-Means++ [CITE]. In Random Centers, $K$
-samples are randomly chosen from the dataset to serve as the initial cluster
-centers. In K-Means++, $1$ sample is randomly chosen from the dataset to serve
-as the initial cluster center. Then, $K-1$ samples are iteratively selected,
-where the probability of a sample being chosen as a cluster center is the
-distance of the sample from previously chosen centers [CITE]. The K-Means
-algorithm is susceptible to initial starting cluster centers, where different
-initial values can result in a different classification.
+standard methods: Random Centers, and K-Means++ \cite{Arthur2007}. In Random
+Centers, $K$ samples are randomly chosen from the dataset to serve as the
+initial cluster centers. In K-Means++, $1$ sample is randomly chosen from the
+dataset to serve as the initial cluster center. Then, $K-1$ samples are
+iteratively selected, where the probability of a sample being chosen as a
+cluster center is the distance of the sample from previously chosen centers
+\cite{Arthur2007}. The K-Means algorithm is susceptible to initial starting
+cluster centers, where different initial values can result in a different
+classification \cite{Franti2018}.
 
@@
 ISODATA is another frequently used clustering algorithm and is based on the
-K-Means algorithm. Like K-Means, cluster centers are iteratively updated and
+K-Means algorithm \cite{Ball1965}. Like K-Means, cluster centers are iteratively updated and
 reassigned to the samples in the dataset. However, ISODATA determines the number
@@
 quality of clusters produced and the number of iterations required to converge
-      [CITE].
+\cite{Franti2018}.
 
@@
 \begin{itemize}
-      \item \textbf{Step 1:} Choose $1$ observation as a cluster center from the
+      \item \textbf{Step 1:} Choose one observation as a cluster center from the
             dataset.
@@
             each dimension. Then, split each cluster along the dimension with
-            the greatest standard deviation.
+            the greatest standard deviation and move to \textbf{Step 2}.
 \end{itemize}
@@
 % =========================================================================== %
-
-\begin{table}[ht]
-      \begin{tabular}{ |p{1.25cm}||p{1.25cm} p{1.25cm} p{1.25cm} p{1.25cm} p{1.25cm} p{1.25cm}|}
-            \hline
-            \multicolumn{7}{|c|}{\textbf{ISODATA}}                               \\
-            \hline
-            Init. & Iter.  & Clusters & ARI             & RI    & MI    & HI     \\
-            \hline
-            \multicolumn{7}{|c|}{\textit{Iris}}                                  \\[1ex]
-            ISODS & 13.46  & 2.599    & \textbf{0.646}  & 0.826 & 0.173 & 0.653  \\
-            KM++  & 15.43  & 2.544    & 0.628           & 0.816 & 0.184 & 0.631  \\
-            Rand  & 11.90  & 2.41     & 0.607           & 0.802 & 0.198 & 0.604  \\
-            \hline
-            \multicolumn{7}{|c|}{\textit{Glass (Normalized)}}                    \\[1ex]
-            ISODS & 3.19   & 1.066    & \textbf{0.0018} & 0.266 & 0.734 & -0.468 \\
-            KM++  & 4.02   & 1.028    & 0.0011          & 0.265 & 0.735 & -0.470 \\
-            Rand  & 3.69   & 1.055    & 0.0013          & 0.266 & 0.733 & -0.467 \\
-            \hline
-            \multicolumn{7}{|c|}{\textit{Wine}}                                  \\[1ex]
-            ISODS & 9.16   & 1.836    & \textbf{0.313}  & 0.618 & 0.382 & 0.236  \\
-            KM++  & 8.76   & 1.816    & 0.307           & 0.612 & 0.388 & 0.224  \\
-            Rand  & 9.26   & 1.810    & 0.304           & 0.608 & 0.391 & 0.217  \\
-            \hline 
-            \multicolumn{7}{|c|}{\textbf{K-Means}}                               \\
-            \hline
-            \multicolumn{7}{|c|}{\textit{Iris}}                                  \\[1ex]
-            ISODS & 2.29   & 3        & \textbf{0.623}  & 0.833 & 0.167 & 0.667  \\
-            KM++  & 4.149  & 3        & 0.584           & 0.811 & 0.189 & 0.621  \\
-            Rand  & 29.402 & 3        & 0.569           & 0.797 & 0.203 & 0.593  \\
-            \hline
-            \multicolumn{7}{|c|}{\textit{Glass (Normalized)}}                    \\[1ex]
-            ISODS & 5.651  & 6        & 0.164           & 0.665 & 0.335 & 0.329  \\
-            KM++  & 4.866  & 6        & \textbf{0.173}  & 0.643 & 0.357 & 0.285  \\
-            Rand  & 77.178 & 6        & 0.151           & 0.634 & 0.366 & 0.268  \\
-            \hline
-            \multicolumn{7}{|c|}{\textit{Wine}}                                  \\[1ex]
-            ISODS & 3.145  & 3        & \textbf{0.882}  & 0.947 & 0.053 & 0.895  \\
-            KM++  & 5.388  & 3        & 0.839           & 0.926 & 0.074 & 0.851  \\
-            Rand  & 24.775 & 3        & 0.834           & 0.920 & 0.080 & 0.840  \\
-            \hline
-      \end{tabular}
-      \caption{UCI benchmark dataset results averaged over 1000 iterations of
-            ISODATA and K-Means comparing three different initial cluster center
-            algorithms (\textit{Init.}). Results in the table contain
-            information on average iterations required (\textit{Iter.}), average
-            number of clusters created (\textit{Clusters}), and validity
-            measures.}
-      \label{tab:ucitable}
-\end{table}
 
@@
 two sets of data to determine agreement and disagreement \cite{Rezaei2016}. The
-pair-counting measures applied are \textit{rand index} (RI - agreement
-probability), \textit{adjusted rand index} (ARI), \textit{mirkin's index} (MI -
-disagreement probability), and \textit{hubert's index} (HI - $(P(agree)$ -
+pair-counting measures applied are \textit{Rand index} (RI - agreement
+probability), \textit{adjusted Rand index} (ARI), \textit{Mirkin's index} (MI -
+disagreement probability), and \textit{Hubert's index} (HI - $(P(agree)$ -
 $P(disagree))$) \cite{Hubert1985, Meila2003}. When comparing ground truth labels
@@
 
-\begin{figure}[ht]
-      \centering
-      \begin{subfigure}[b]{0.4\textwidth}
-            \centering
-            \includegraphics[width=\linewidth]{salinasgt.png}
-            \caption{Salinas-A Ground Truth}
-            \label{fig:salinasgt}
-      \end{subfigure}
-      \begin{subfigure}[b]{0.4\textwidth}
-            \centering
-            \includegraphics[width=\linewidth]{risodata.png}
-            \caption{Random Centers (ARI=0.797)}
-            \label{fig:salinasr}
-      \end{subfigure}\hfill
-      \centering
-      \begin{subfigure}[b]{0.4\textwidth}
-            \centering
-            \includegraphics[width=\linewidth]{kisodata.png}
-            \caption{KM++ Centers (ARI=0.781)}
-            \label{fig:salinask}
-      \end{subfigure}
-      \begin{subfigure}[b]{0.4\textwidth}
-            \centering
-            \includegraphics[width=\linewidth]{iisodata.png}
-            \caption{ISODS Centers (ARI=0.758)}
-            \label{fig:salinasi}
-      \end{subfigure}\hfill \caption{Heatmap showing cluster assignments of each
-            pixel using ISODATA with different initial cluster center methods.}
-      \label{fig:salinas}
-\end{figure}
+The ISODATA algorithm does not have a determined number of clusters to generate.
+Instead, the user sets the max clusters desired, and the algorithm will generate
+a number of clusters between and the max and half the max number desired. Since
+the output depends on the data, the ISODATA parameters, and the initial centers,
+we include a column to depict the average number of clusters.
+
+\begin{table}[ht]
+      \caption{UCI benchmark dataset results averaged over 1000 iterations of
+            ISODATA comparing three different initial cluster center algorithms
+            (\textit{Init.}). ARI - adjusted Rand index; RI - Rand index; MI -
+            Mirkin's index; HI - Hubert's index.}
+      \begin{tabular}{ |p{1.25cm}||p{1.25cm} p{1.25cm} p{1.25cm} p{1.25cm} p{1.25cm} p{1.25cm}|}
+            \hline
+            \multicolumn{7}{|c|}{\textbf{ISODATA}}                               \\
+            \hline
+            Init. & Iter.  & Clusters & ARI             & RI    & MI    & HI     \\
+            \hline
+            \multicolumn{7}{|c|}{\textit{Iris}}                                  \\
+            \hline
+            ISODS & 13.46  & 2.599    & \textbf{0.646}  & 0.826 & 0.173 & 0.653  \\
+            KM++  & 15.43  & 2.544    & 0.628           & 0.816 & 0.184 & 0.631  \\
+            Rand  & 11.90  & 2.41     & 0.607           & 0.802 & 0.198 & 0.604  \\
+            \hline
+            \multicolumn{7}{|c|}{\textit{Glass (Normalized)}}                    \\
+            \hline
+            ISODS & 3.19   & 1.066    & \textbf{0.0018} & 0.266 & 0.734 & -0.468 \\
+            KM++  & 4.02   & 1.028    & 0.0011          & 0.265 & 0.735 & -0.470 \\
+            Rand  & 3.69   & 1.055    & 0.0013          & 0.266 & 0.733 & -0.467 \\
+            \hline
+            \multicolumn{7}{|c|}{\textit{Wine}}                                  \\
+            \hline
+            ISODS & 9.16   & 1.836    & \textbf{0.313}  & 0.618 & 0.382 & 0.236  \\
+            KM++  & 8.76   & 1.816    & 0.307           & 0.612 & 0.388 & 0.224  \\
+            Rand  & 9.26   & 1.810    & 0.304           & 0.608 & 0.391 & 0.217  \\
+            \hline 
+            \multicolumn{7}{|c|}{\textbf{K-Means}}                               \\
+            \hline
+            \multicolumn{7}{|c|}{\textit{Iris}}                                  \\
+            \hline
+            ISODS & 2.29   & 3        & \textbf{0.623}  & 0.833 & 0.167 & 0.667  \\
+            KM++  & 4.149  & 3        & 0.584           & 0.811 & 0.189 & 0.621  \\
+            Rand  & 29.402 & 3        & 0.569           & 0.797 & 0.203 & 0.593  \\
+            \hline
+            \multicolumn{7}{|c|}{\textit{Glass (Normalized)}}                    \\
+            \hline
+            ISODS & 5.651  & 6        & 0.164           & 0.665 & 0.335 & 0.329  \\
+            KM++  & 4.866  & 6        & \textbf{0.173}  & 0.643 & 0.357 & 0.285  \\
+            Rand  & 77.178 & 6        & 0.151           & 0.634 & 0.366 & 0.268  \\
+            \hline
+            \multicolumn{7}{|c|}{\textit{Wine}}                                  \\
+            \hline
+            ISODS & 3.145  & 3        & \textbf{0.882}  & 0.947 & 0.053 & 0.895  \\
+            KM++  & 5.388  & 3        & 0.839           & 0.926 & 0.074 & 0.851  \\
+            Rand  & 24.775 & 3        & 0.834           & 0.920 & 0.080 & 0.840  \\
+            \hline
+      \end{tabular}
+      \label{tab:ucitable}
+\end{table}
 
@@
 \begin{table}[ht]
+      \caption{Salinas-A dataset average and best results over 1000 iterations
+            of ISODATA comparing three different initial cluster center
+            algorithms (\textit{Init.}).}
       \begin{tabular}{ |p{1.25cm}||p{1.25cm} p{1.25cm} p{1.25cm} p{1.25cm} p{1.25cm} p{1.25cm}|}
@@
             \hline
-            \multicolumn{7}{|c|}{\textit{Average}}                             \\[1ex]
+            \multicolumn{7}{|c|}{\textit{Average}}                             \\
+            \hline
             ISODS & 2.772 & 2.42     & \textbf{0.156} & 0.388 & 0.612 & -0.224 \\
@@
             \hline
-            \multicolumn{7}{|c|}{\textit{Best}}                                \\[1ex]
+            \multicolumn{7}{|c|}{\textit{Best}}                                \\
+            \hline
             ISODS & 2     & 6        & 0.758          & 0.921 & 0.078 & 0.842  \\
@@
       \end{tabular}
-      \caption{Salinas-A dataset average and best results over 1000 iterations
-            of ISODATA comparing three different initial cluster center
-            algorithms (\textit{Init.}). Results in the table contain
-            information on iterations required (\textit{Iter.}), number of
-            clusters created (\textit{Clusters}), and validity measures.}
       \label{tab:salinastable}
 \end{table}
+
+\begin{figure}[ht]
+      \centering
+      \begin{subfigure}[b]{0.24\textwidth}
+            \centering
+            \includegraphics[width=\linewidth]{salinasgt.png}
+            \label{fig:salinasgt}
+      \end{subfigure}
+      \begin{subfigure}[b]{0.24\textwidth}
+            \centering
+            \includegraphics[width=\linewidth]{risodata.png}
+            \label{fig:salinasr}
+      \end{subfigure}
+      \centering
+      \begin{subfigure}[b]{0.24\textwidth}
+            \centering
+            \includegraphics[width=\linewidth]{kisodata.png}
+            \label{fig:salinask}
+      \end{subfigure}
+      \begin{subfigure}[b]{0.24\textwidth}
+            \centering
+            \includegraphics[width=\linewidth]{iisodata.png}
+            \label{fig:salinasi}
+      \end{subfigure}\hfill \caption{Heatmap showing cluster assignments of each
+            pixel using ISODATA with different initial cluster center methods.
+            (a) Salinas-A Ground Truth; (b) Random Centers (ARI=0.797); KM++
+            Centers (ARI=0.781); ISODS Centers (ARI=0.758).}
+      \label{fig:salinas}
+\end{figure}
 
@@
 This study presents a method for generating initial cluster centers based on a
-splitting approach, ISODS. We tested ISODS on bechmark datasets as well as the
-Salinas-A hyperspectral image dataset. Both tests on average shows the
-performance of ISODS using external cluster validation measures, especially on
-the \textit{adjusted rand index} measure. Additonally, ISODS may decrease the
-number of iterations required for a clustering algorithm such as ISODATA or
-K-Means to converge. However, due to the complexity of the ISODS approach,
-specifically calculating the means of the clusters, the number of iterations
-taken by ISODS should be considered. When used in conjunction with ISODATA,
-ISODS also shows an average number of clusters generated closer to the expected
+splitting approach, Iterative Self-Organizing Data Split (ISODS). We tested ISODS on
+bechmark datasets as well as the Salinas-A hyperspectral image dataset. Both
+tests on average shows the performance of ISODS using external cluster
+validation measures, especially on the \textit{adjusted Rand index} measure.
+Additonally, ISODS may decrease the number of iterations required for a
+clustering algorithm such as ISODATA or K-Means to converge. However, due to the
+complexity of the ISODS approach compared to other initial centroid algorithms,
+specifically calculating the means of the clusters, the context should be
+considered when applying this method. When used in conjunction with ISODATA,
+ISODS shows an average number of clusters generated closer to the expected
 output.

### 2021-01-11 paper/paper.tex
add hyper links

@@
 \usepackage{subcaption}
+\usepackage[hidelinks]{hyperref}
 
@@
 \title{\bf ISODS: An ISODATA-Based Initial Centroid Algorithm}
-\author{Noah Gardner$^1$, Sahar Yarmohammadtoosky$^2$, \\
-      Anthony Phan$^3$, and Chih-Cheng Hung$^4$\\[1ex]
+\author{
+      Noah Gardner$^1$,
+      Sahar Yarmohammadtoosky$^2$, \\
+      Anthony Phan$^3$,
+      and Chih-Cheng Hung$^4$\\[1ex]
       \sl \small Kennesaw State University, Marietta, GA, USA$^{1, 3, 4}$\\
@@
       Iran$^2$\\[1ex]
-      \small $^1$ngardn10@students.kennesaw.edu, \small
-      $^2$yarmohamadishr@gmail.com, \\
-      \small $^3$aphan5@students.kennesaw.edu, \small $^4$chung1@kennesaw.edu,
-      \\}
+      \small \href{mailto:ngardn10@students.kennesaw.edu}{$^1$ngardn10@students.kennesaw.edu},
+      \small \href{mailto:yarmohamadishr@gmail.com}{$^2$yarmohamadishr@gmail.com}, \\
+      \small \href{mailto:aphan5@students.kennesaw.edu}{$^3$aphan5@students.kennesaw.edu},
+      \small \href{mailto:chung1@kennesaw.edu}{$^4$chung1@kennesaw.edu} \\
+}
 \date{}
@@
 beginning of the original process. Although more complex than other methods of
-generating initial cluster centers, this approach guarantees variance between
+generating initial cluster centers, this approach guarantees separation between
 clusters. As a result, ISODS can be used to generate clusters for other
@@
             of ISODATA comparing three different initial cluster center
-            algorithms (\textit{Init.}).}
+            algorithms (\textit{Init.}). ARI - adjusted Rand index; RI - Rand
+            index; MI - Mirkin's index; HI - Hubert's index.}
       \begin{tabular}{ |p{1.25cm}||p{1.25cm} p{1.25cm} p{1.25cm} p{1.25cm} p{1.25cm} p{1.25cm}|}
@@
             \includegraphics[width=\linewidth]{salinasgt.png}
+            \caption{}
             \label{fig:salinasgt}
@@
             \includegraphics[width=\linewidth]{risodata.png}
+            \caption{}
             \label{fig:salinasr}
@@
             \includegraphics[width=\linewidth]{kisodata.png}
+            \caption{}
             \label{fig:salinask}
@@
             \includegraphics[width=\linewidth]{iisodata.png}
+            \caption{}
             \label{fig:salinasi}
@@
             pixel using ISODATA with different initial cluster center methods.
-            (a) Salinas-A Ground Truth; (b) Random Centers (ARI=0.797); KM++
+            (a) Salinas-A Ground Truth; (b) Random Centers (ARI=0.798); KM++
             Centers (ARI=0.781); ISODS Centers (ARI=0.758).}
